The error in Bayesian phylogenetic inference when speciation is protracted

### 2019-05-08 content.tex
Travis uses Xenial

  % One or two sentences providing a basic
  % introduction to the field,
  % comprehensible to a scientist in any discipline.
  The tools for reconstructing phylogenetic relationships between taxonomic 
  units (e.g. species) have become very advanced in the last three decades. 
    % Two to three sentences of
  % more detailed background, comprehensible to
  % scientists in related disciplines.
  Among the most popular tools are Bayesian approaches, 
  such as \\ BEAST, MrBayes and RevBayes, 
  that use efficient tree sampling routines to create a posterior probability distribution 
  of the phylogenetic tree. 
  A feature of these approaches is the possibility to incorporate known or hypothesized structure of the phylogenetic tree through the tree prior. 
  It has been shown that the effect of the prior on the posterior distribution 
  of trees can be substantial.

@@

The model priors in Bayesian phylogenetic reconstruction 
can be grouped into three categories: (1) site model, specifying 
nucleotide substitutions, (2) clock model, specifying
the rate of mutation per lineage in time, and (3) tree model, 
constituting the speciation model underlying branching events (speciation) 
and branch termination (extinction).
The choice of site model (\cite{posada_and_buckley_2004}), 
clock model (\cite{baele_et_al_2012}) 
or tree prior (\cite{moller2018, yang_and_ranalla_2005}) is known to affect
the posterior.

@@

\richel{
  Note there is no twinning, as in the razzo package.
  Already sent an email to discuss this with RSE.
}

As described above, per alignment, we do four different Bayian analyses,
as we use two different site models and two different clock models.
We know the generative site model (which is JC69) and clock 
model (a strict clock), but want to explore how often the correct model
is indeed preferred. Per alignment, we measure the estimated marginal 
likelihood (which is the probability of the data given the model)
for each of the four models and measure their relative propertions.
To estimate the marginal likelihood, we use the novel 
Nested Sampling approach (\cite{maturana_et_al_2018}),
which we configure to have an effective sample size (in estimating
the marginal likelihood) of at least 200, by using a single-particle
sub-chain with a length equal to the MCMC interval.
\richel{
  There are no rules described in the setup of the NS
  (for example, what should the relation between MCMC chain length
  and NS sub-chain length be?),
  but one can experiment with a sub-chain length and number of particles.
  I will need to do some pilot studies to verify if this is
  the correct way to do and describe this.  
}

@@

In this article, we showcase the effect of sampling
and the certainty when selecting a (site and clock) model. 
We show the effect sampling has on the inference error,
for the nLTT distributions generated from assuming the (correct) Jukes-Cantor 
site model and a strict clock model. We do so per parameter combination,
displaying the distribution as a violin plot.
In the supplementary information, we display these error distributions 
of the general data set all combined, or separated per site or clock model.
The certainty in the model selection is again shown per parameter setting,
as a stacked bar. 
\richel{
  I am unsure how to depict that stacked bar and its uncertainty.
  Sure, the stacked bars (without uncertainty) will work,
  but information about uncertainty is lost. 
  See figure 'Showing a stacked bar with uncertainty'
  for three equivalent ways to display a stacked bar with uncertainty.
}

@@

\richel{put this section here, as the journal does not request for this}
We would like to thank the Center for Information Technology of the University of Groningen for their support
and for providing access to the Peregrine high performance computing cluster.

@@

\richel{put this section here, as the journal does not request for this}
RSE conceived the idea for this experiment. 
RJCB created and tested the experiment, and wrote the first draft of the manuscript. 
RSE contributed substantially to revisions.

>>>>

  % One or two sentences providing a basic
  % introduction to the field,
  % comprehensible to a scientist in any discipline.
  The tools for reconstructing phylogenetic relationships between taxonomic 
  units (e.g. species) have become very advanced in the last three decades. 
    % Two to three sentences of
  % more detailed background, comprehensible to
  % scientists in related disciplines.
  Among the most popular tools are Bayesian approaches, 
  such as BEAST, MrBayes and RevBayes, 
  that use efficient tree sampling routines to create a 
  posterior probability distribution 
  of the phylogenetic tree. 
  A feature of these approaches is the possibility to incorporate known 
  or hypothesized structure of the phylogenetic tree through the tree prior. 
  It has been shown that the effect of the prior on the posterior distribution 
  of trees can be substantial.

@@

The model priors in Bayesian phylogenetic reconstruction 
can be grouped into three categories: (1) site model, specifying 
nucleotide substitutions, (2) clock model, specifying
the rate of mutation per lineage in time, and (3) tree model, 
constituting the speciation model underlying branching events (speciation) 
and branch termination (extinction).
The choice of site model (\cite{posada_and_buckley_2004}), 
clock model (\cite{baele_et_al_2012}) 
or tree prior (\cite{moller2018, yang_and_ranalla_2005}) is known to affect
the posterior. There is evidence, however, that the 
tree prior and molecular clock do not do so substantially affect 
the estimation of diversification rates (\cite{sarver2019choice})).

@@

TWINNING

As described above, per alignment, we do four different Bayian analyses,
as we use two different site models and two different clock models.
We know the generative site model (which is JC69) and clock 
model (a strict clock), but want to explore how often the correct model
is indeed preferred. Per alignment, we measure the estimated marginal 
likelihood (which is the probability of the data given the model)
for each of the four models and measure their relative propertions.
To estimate the marginal likelihood, we use the novel 
Nested Sampling approach (\cite{maturana_et_al_2018}),
which we configure to have a relative error $\epsilon$ of $1^{-12}$.

@@

In this article, we showcase the effect of sampling
and the certainty when selecting a (site and clock) model. 
We show the effect sampling has on the inference error,
for the nLTT distributions generated from assuming the (correct) Jukes-Cantor 
site model and a strict clock model. We do so per parameter combination,
displaying the distribution as a violin plot.
In the supplementary information, we display these error distributions 
of the general data set all combined, or separated per site or clock model.
The certainty in the model selection is again shown per parameter setting,
as a stacked bar.

HIERO: DESCRIBE PLOTTING.

@@

We would like to thank the Center for Information Technology of the University of Groningen for their support
and for providing access to the Peregrine high performance computing cluster.

@@

RSE conceived the idea for this experiment. 
RJCB created and tested the experiment, and wrote the first draft of the manuscript. 
RSE contributed substantially to revisions.
